\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-10-07 submission/manuscript.Rmd
update

While the delay of stimulus re-exposure stimulus did not have a significant effect on participants' beliefs of reality ($OR = 1.00$, $95\CI = [0.99, 1.00]$ ), judgement confidence was found to be negatively associated with re-exposure delay when the faces were judged as real ($\beta = -0.006$, $95\%~CI = [-0.1, 0.002]$, $p = .004$)

@@

Attractiveness had a significant positive and linear relationship ($R^2_{marginal}$ = 2.0\%) with the belief that a stimulus was real ($\beta_{poly1} = 16.57$, $95\%~CI = [7.33, 25.82]$, $z = 3.51$, $p < .001$) for males, and a quadratic relationship for females ($\beta_{poly2} = 7.82$, $95\%~CI = [1.81, 13.84]$, $z = 2.55$, $p = .011$), with both non-attractive and attractive faces being judged as more real. Attractiveness was also found to have a significant positive and quadratic relationship with confidence in judging faces both as real ($\beta_{poly2} = 4.30, $, $95\%~CI = [0.97, 7.64]$, $z = 2.53$, $p = .011$) and as fake ($\beta_{poly2} = 5.23, $, $95\%~CI = [0. 86, 9.60]$, $z = 2.35$, $p = .019$) for females. For males, however, a significant negative and quadratic relationship was found between attractiveness ratings and belief confidence only for faces judged as fake ($\beta_{poly2} = -9.92, $, $95\%~CI = [-18.99, -0.86]$, $z = -2.15$, $p = .032$). There was no interaction with reported self-attractiveness.

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that the stimulus is artificially generated. The most striking result, in our opinion, is that despite all the stimuli being real faces from the same database, all participants, when given the information, believed (to high degrees of confidence) that a significant proportion of them were fake. This finding could thus be seen as a testimony to both the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted for less than 20\% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

<!-- Attractiveness: males -->
Interestingly, we found a significant positive linear relationship in male participants for both attractiveness and beauty on simulation monitoring that we could interpret under an evolutionary lens. Specifically, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial; @buunk2002age]. However, while it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces for males, rendering them significantly more salient than unattractive faces and in turn distorted the relationship with simulation monitoring, future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or physiological markers [in particular, heart rate deceleration, @skora2022functional].

@@

<!-- trustworthiness -->
Our results found a positive linear trend between trustworthiness and simulation monitoring for females only; suggesting trustworthiness ratings increased their beliefs in realness. Given prior evidence that faces presented as computer-generated were rated less trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a linear association to be more clearly present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. However, if that was the case, faces judged as highly untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real, leading to a quadratic relationship between trustworthiness and simulation monitoring instead. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. Interestingly, there were significant linear relationships between familiarity and confidence judgements for males only, where highly familiar faces were rated as real and unfamiliar faces as fake with greater confidence. This could be taken into perspective with previous studies, which report females to be relatively superior at face recall and recognition [@mishra2019gender; @lewin2002sex], displaying higher levels of facial memory [@sommer2013sex] and exhibiting greater metacognitive awareness [@kaplan2012effect]. Hence, it is possible that the confidence in evaluating faces are less malleable to external factors such as face familiarity for females. Indeed, metacognitive awareness has been associated with overconfidence in making judgements [@mata2013metacognitive; @klein2018metacognitive]. However, it should be noted that this has not been corroborated by recent research, which found males were significantly overconfident in face recall, regardless of their performance accuracy [@bailey2021gender; @herbst2020gender]. Furthermore, the present study's distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. As such, future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments. Notably, greater narcissistic tendencies in dimensions such as acclaim seeking and manipulativeness was found to be associated with a greater belief that the faces presented were real. Indeed, this finding is in line with recent research suggesting that narcissists are more susceptible to fake news [@piksa2022cognitive; @sindermann2020short], possibly because they were less likely to engage in cognitive strategies such as reflective thinking [@littrell2020overconfidently; @ahadzadeh2021social] and are more vigilant when attending to external stimuli [@carolan2017searching; @grapsas2020and; @eddy2021self]. As such, individuals with greater narcissistic tendencies could have perceived the face images as more salient and engaged in less analytical reasoning, leading to their increased perceptions of the faces being real.

@@

<!-- for thrill-seeking, a tolerance of cognitive uncertainty (note that the intolerance to uncertainty scale, which largely measures the anxiety associated with uncertain events, did not show any link) -->

@@

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->
Unsurprisingly, we found that individuals with more persecutory ideations (a form of paranoia where one believes others intend to cause them harm, @buck2019capturing) tend to rate the faces presented as real relative to fake. Given previous reports that people with higher levels of paranoia are more sensitive to cues of social threat [@fornells2015people; @freeman2003can], it could thus be possible that people with more persecutory ideations perceived the faces with greater salience and emotionality, hence increasing their beliefs of its realness. Indeed, this is in line with previous findings that persecutory delusions are predicted by a greater sense of presence in a VR environment populated with virtual characters [@freeman2005psychology].

In contrast, however, intolerance to uncertainty, a trait related to enhanced threat appraisal, was found to be negatively associated with confidence in real judgements [@chen2016intolerance; @jensen2014intolerance]. This is consistent with previous reports that uncertainty intolerance correlates with under-confidence in decision-making in ambiguous scenarios [@wei2021confidence; @jensen2014intolerance]. While the findings seem contradictory, they could possibly be interpreted through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a means to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. In fact, previous studies report that individuals with greater paranoia and anxiety engaged in more affective processing of external ambiguous stimuli [@fornells2015people]. This is also in line with prior findings which suggest that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2016phenomenal]. As such, it is possible that individuals high in such anxiety-related traits are more sensitive to social cues and perceived the ambiguous face images with greater emotionality, therefore having to engage in emotional regulation strategies that attenuated their confidence in real judgements. Future studies should clarify the role of emotions, particularly those related to anxiety, and attempt at delineating processes related to emotion perception and reaction from emotion regulation.

@@

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, participants were more confident in their judgements for faces perceived as real following a shorter re-exposure delay. This may be because shorter re-exposure delays led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. Indeed, this finding is in line with studies that show an up-regulation of emotions when fictional stimuli were associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. However, that being the case, we expect shorter re-exposure delays to be significantly related to a greater belief of the faces being real. Future studies should further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

<!--a shorter re-exposure delay was associated with an increased likelihood of judging the faces as real.  Although the distribution variance in re-exposure delay is small, recent research using event-related brain potentials report representations of faces to be qualitatively similar regardless of the degree and type of familiarity (e.g., personal or media-based), only distinguishing between familiar and unfamiliar faces [@wiese2021familiarity]. Moreover, fictional stimuli has been shown to up-regulate emotions when associated with familiarity [@sperduti2016paradox; @makowski2017being]. -->

In conclusion, research on our ability to discriminate between what is real from what is not is increasingly pertinent, given the current expansion on technologies related to virtual and augmented realities. The aim of the present study was to examine whether facial attractiveness significantly influences our simulation monitoring decisions. In line with our hypothesis, we found faces rated as attractive and unattractive to be perceived as more real for females as a result of an increased salience. In contrast, the observation of a positive linear relationship between attractiveness and beauty with simulation monitoring decisions, adjusted for evolutionarily advantageous features such as trustworthiness, suggest a more complex relationship than hypothesized for males. As such, future studies could take greater consideration of the multidimensionality of facial attractiveness in investigating its role in simulation monitoring decisions. Furthermore, it is important to note that besides the traits of narcissistic acclaim-seeking and manipulativeness, and the persecutory ideation dimension of paranoid thinking, all of the aforementioned dispositional traits did only influence the confidence judgments, but not the simulation monitoring bias *per se*. Hence, future studies could expand upon the present study to further investigate the types of and degree to which different inter-individual variables influence our beliefs of reality.

>>>>

While the delay of stimulus re-exposure stimulus did not have a significant effect on participants' beliefs of reality ($OR = 1.00$, $95\%~CI = [0.99, 1.00]$ ), judgement confidence was found to be negatively associated with re-exposure delay when the faces were judged as real ($\beta = -0.006$, $95\%~CI = [-0.1, 0.002]$, $p = .004$)

@@

Attractiveness had a significant positive and linear relationship ($R^2_{marginal}$ = 2.0\%) with the belief that a stimulus was real ($\beta_{poly1} = 16.57$, $95\%~CI = [7.33, 25.82]$, $z = 3.51$, $p < .001$) for males, and a quadratic relationship for females ($\beta_{poly2} = 7.82$, $95\%~CI = [1.81, 13.84]$, $z = 2.55$, $p = .011$), with both non-attractive and attractive faces being judged as more real. Attractiveness was also found to have a significant positive and quadratic relationship with confidence in judging faces both as real ($\beta_{poly2} = 4.30$, $95\%~CI = [0.97, 7.64]$, $z = 2.53$, $p = .011$) and as fake ($\beta_{poly2} = 5.23$, $95\%~CI = [0. 86, 9.60]$, $z = 2.35$, $p = .019$) for females. For males, however, a significant negative and quadratic relationship was found between attractiveness ratings and belief confidence only for faces judged as fake ($\beta_{poly2} = -9.92$, $95\%~CI = [-18.99, -0.86]$, $z = -2.15$, $p = .032$). There was no interaction with reported self-attractiveness.

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that a stimulus was artificially generated. The most striking result, in our opinion, is that despite all the stimuli being real faces from the same database, all participants believed (to high degrees of confidence) that a significant proportion of them were fake. This finding can be seen as a testimony not only to the effectiveness of our instructions, nut also to the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted together for less than 20\% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

<!-- Attractiveness: males -->
Interestingly, we found a significant positive linear relationship in male participants for both attractiveness and beauty on simulation monitoring that we could interpret under an evolutionary lens. Specifically, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial; @buunk2002age]. It is thus possible that the evolutionary weight associated with attractiveness skewed men's perceived saliency towards attractive faces, rendering them significantly more salient than unattractive faces and in turn distorted the relationship with simulation monitoring. However, future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or physiological markers [e.g., heart rate deceleration, @skora2022functional].

@@

<!-- trustworthiness -->
Our results found a positive linear trend between trustworthiness and simulation monitoring for females only. Given prior evidence that faces presented as computer-generated were rated less trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a linear association to be more clearly present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could be the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. However, if that was the case, faces judged as highly untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real, leading to a quadratic relationship between trustworthiness and simulation monitoring instead. Further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. Interestingly, there were significant linear relationships between familiarity and confidence judgements for males only, where familiarity increased the confidence of reality beliefs. This could be taken into perspective with previous studies, which report females to be relatively superior at face recall and recognition [@mishra2019gender; @lewin2002sex], displaying higher levels of facial memory [@sommer2013sex] and exhibiting greater metacognitive awareness [@kaplan2012effect]. Hence, it is possible that the confidence in evaluating faces are less malleable to external factors such as face familiarity for females. Indeed, metacognitive awareness has been associated with overconfidence in making judgements [@mata2013metacognitive; @klein2018metacognitive]. However, it should be noted that this has not been corroborated by recent research, which found males were significantly overconfident in face recall, regardless of their performance accuracy [@bailey2021gender; @herbst2020gender]. Furthermore, the present study's distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. As such, future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments. Notably, greater narcissistic tendencies in dimensions such as acclaim seeking and manipulativeness were found to be associated with a higher number of faces judged as real. Indeed, this finding is in line with recent research suggesting that narcissists are more susceptible to fake news [@piksa2022cognitive; @sindermann2020short], possibly because they were less likely to engage in cognitive strategies such as reflective thinking [@littrell2020overconfidently; @ahadzadeh2021social] and are more vigilant when attending to external stimuli [@carolan2017searching; @grapsas2020and; @eddy2021self]. As such, individuals with greater narcissistic tendencies could have perceived the face images as more salient and engaged in less analytical reasoning, leading to their increased perceptions of the faces being real.

@@

@@

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->
Our findings suggest - though with weak significance - a positive link between paranoid ideation and the tendency to believe that the stimuli were real. Given previous reports that people with higher levels of paranoia are more sensitive to cues of social threat [@fornells2015people; @freeman2003can], it could thus be possible that stronger paranoid traits was associated with a greater perception of faces as salient and emotional, hence increasing their beliefs of its realness. This hypothesis, if confirmed by future studies, would be in line with previous findings that persecutory delusions are predicted by a greater sense of presence in a VR environment populated with virtual characters [@freeman2005psychology].

In contrast, intolerance to uncertainty, a trait related to enhanced threat appraisal, was found to be negatively associated with the confidence in real judgements [@chen2016intolerance; @jensen2014intolerance]. This is consistent with previous reports that uncertainty intolerance correlates with under-confidence in decision-making in ambiguous scenarios [@wei2021confidence; @jensen2014intolerance]. **WHY IS THIS CONTRADICTORY? the next part seems to be about emotion regulation. While interesting, it's a bit of a non-sequitur no? And the previous explanation could almost be enough.** While the findings seem contradictory, they could possibly be interpreted through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a means to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. In fact, previous studies report that individuals with greater paranoia and anxiety engaged in more affective processing of external ambiguous stimuli [@fornells2015people]. This is also in line with prior findings which suggest that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2019phenomenal]. As such, it is possible that individuals high in such anxiety-related traits are more sensitive to social cues and perceived the ambiguous face images with greater emotionality, therefore having to engage in emotional regulation strategies that attenuated their confidence in real judgements. Future studies should clarify the role of emotions, particularly those related to anxiety, and attempt at delineating processes related to emotion perception and reaction from emotion regulation.

@@

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, participants were more confident in their judgements for faces perceived as real following a shorter re-exposure delay. This may be because shorter re-exposure delays led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. Indeed, this finding is in line with studies that show an up-regulation of emotions when fictional stimuli were associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. If that was the case, we would expect shorter re-exposure delays to impact the decision bias as well towards reality, rather than simply the confidence. Future studies should further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

<!-- Conclusion -->
In conclusion, understanding the factors driving our beliefs that ambiguous stimuli are real or not is becoming increasingly pertinent, given the current expansion on technologies related to virtual and augmented realities. The aim of the present study was to examine whether facial attractiveness significantly influences our simulation monitoring decisions. We found faces rated as attractive to be perceived as more real, with a possible sexual dimorphism affecting the shape of the relationship. We also found that inter-individual traits, such as narcissistic acclaim-seeking and manipulativeness, as well as persecutory ideation, were related to a systematic bias towards beliefs that the stimuli were real or fake. Limitations include [**TO ADD**].
